Stack Safety as a Security Property

### 2020-07-09 main.tex
light pass over 7; releasing lock

{\em Stack integrity} is a property of a system with regard to a call map
\(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
and \(\mpstate{\mach_e}{\pol_e}\) in every legal trace exhibits trace integrity
on the prefix up to its return, if any. For any initial state \(\mach_0\)
we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow
\MPT\). For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\), 
\(\propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
system as a whole enjoys stack integrity with regard to \(\callmap\).

@@

{\em Trace confidentiality} will be a property of two traces, \(\MPT\) and
\(\machT\), with regard to a return predicate \(R \subset \MACHS\) that
reflects the states that are considered to have returned. We abstract over
\(R\) because what it means to return depends on the state from which
\(f\) was called. There are three cases in which \(\MPT\) and \(\machT\)
enjoy trace confidentiality with respect to \(R\).

@@

If, for all such pairs of traces, \(\propdef{TraceConfidentialiy} ~
\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
stack confidentiality with regard to \(\callmap\).

@@

  \item
%
    If the machine takes a \emph{call step} and $\contour$ is the contour at
    the call then,
%
    (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
    satisfied,
%
    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
    is satisfied with regard to an arbitrary $\contour$-variant
    of $\mach$,
%
    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
    $\contour$-variant of the machine,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

@@

    \paragraph{Rolling Back to Idealized State}
      Observable integrity will weaken eager integrity by allowing a returned
      state to differ from its call state at high integrity components,
      provided the observable behavior of the induced trace from the return
      matches that of an idealized state whose high integrity components do
      match the call state.

      Given a contour, a call state, and a returned state, we define a
      ``rollback'' function \(\mathit{roll} : \CONTOURS \times \MACHS \times
      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
      the call state on those components that are high integrity in its
      contour, and matches the return state on low integrity components:
      \[\mathit{roll}(\contour,\mach_c,\mach_r)(\component) =
      \begin{cases}
        \mach_c(\component) & \contour(\component) = (\HIGHINT,\_) \\
        \mach_r(\component) & \contour(\component) = (\LOWINT,\_) \\
      \end{cases}\]

@@

@@

    \definition{Observable Trace Integrity}

      An MP-Trace enjoys {\em observable trace integrity} with respect to
      some contour \(\contour\) if its final state (if any) has observable
      behavior that prefixes that of its idealized version, constructed by
      \(\mathit{roll}\):

@@

    For out example, we return to the violation of trace integrity in
    \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
    Whether observable integrity holds depends on the continuing trace, shown
    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
    \(\mach_r\), and the rolled-back state, \(\mach_r' =
    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
    components are restored to their values as of \(\mach_e\). Specifically,
    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
    it prints 0. So observable trace integrity does not hold.

    However, if {\tt g} had an additional instruction to overwrite {\tt z}
    before printing, observable trace integrity would hold.

    \definition{Observable Stack Integrity}

      A system enjoys {\em observable stack integrity} with respect to
      \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
      and contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).

@@

      Trace confidentiality is a property of two traces whose initial machine
      states are variations, with two sub-properties. First, the behavioral
      condition -- that their observable behaviors prior to the end of the
      trace are equivalent -- reflects the intuition that secret values should
      not influence visible behavior. Second, we have the end condition that
      their final states (if any) are equal up to the original variation. The
      end condition over-approximates the extension of the behavioral condition
      to future execution.

      We introduce {\em observable trace confidentiality}, which extends the
      behavioral condition to apply to execution from the final state on.
      This requires restoring varied components to their original values,
      since after return they will once again be accessible legitimately.

@@

    \definition{Observable Trace Confidentiality}

      A pair of traces \(\MPT\) and \(\machT\) starting from
      \(\mpstate{\mach_e}{\pol_e}\) and \(\mach_e'\), respectively, with
      \(\mach_e \approx_\contour \mach_e'\), enjoy observable trace
      confidentiality with respect to \(\contour\) if two conditions hold:

      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
      \pi_o(\machT)\).

      If \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
      \(\mach_r' = \last(\machT)\). Then the induced traces
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
      \hookrightarrow \machT'\) are observationally prefixed,
      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\).

      We write this \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).

@@

      In our example, \cref{fig:conflex} shows how execution continues from
      \(\mach_r'\) and the restored \(\mach_r'' = \mathit{restore}(\contour,
      \mach_e,\mach_e',\mach_r,\mach_r')\). Addresses 108 and 109
      are not restored; every address that was varied and not changed in one
      of the executions is. So 108 contains 43 in the trace from \(\mach_r\),
      and after returning from {\tt g}, {\tt main} prints 85. But in the trace
      from \(\mach_r''\), 108 contains -4, and ultimately 38 is printed.
      So observable trace confidentiality does not hold. This example
      illustrates the weakness of observable confidentiality as a testing
      property; it takes until the very end of the program to determine whether
      a tainted value actually influences observable behavior.

    \definition{Observable Stack Confidentiality}

      From observable trace confidentiality we construct {\em observable stack
      confidentiality}. A system enjoys {\em observable stack confidentiality}
      with respect to \(\callmap\) and \(\codemap\) if for any initial state
      \(\mach_0\) and each transition with entry state
      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).

    \definition{Observable Stack Safety}
      A system enjoys {\em observable stack safety} with respect to a call map
      and a function map if it enjoys both observable stack integrity and
      observable stack confidentiality with respect to them.

>>>>

{\em Stack integrity} is a property of a system with respect to a call map
\(\callmap\) that holds when every \(n\)-transition \(\mpstate{\mach_c}{\pol_c}\)
and \(\mpstate{\mach_e}{\pol_e}\) in every legal trace exhibits trace integrity
on the prefix up to its return, if any. For any initial state \(\mach_0\)
we take the induced trace \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow
\MPT\). For each \(n\)-transition \(\mathit{Tr}_n ~ \MPT ~
\mpstate{\mach_c}{\pol_c} ~ \mpstate{\mach_e}{\pol_e}\) we take the prefix of
the call up to its return, \(\mpstate{\mach_e}{\pol_e} \hookrightarrow
\MPT_{call} | \ret{\mach_c}\). If for any such call trace \(\MPT_{call}\), 
\(\propdef{TraceIntegrity} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\), then the
system as a whole enjoys stack integrity with respect to \(\callmap\).

@@

{\em Trace confidentiality} will be a property of two traces, \(\MPT\) and
\(\machT\), with respect to a return predicate \(R \subset \MACHS\) that
reflects the states that are considered to have returned. We abstract over
\(R\) because what it means to return depends on the state from which
\(f\) was called. There are three cases in which \(\MPT\) and \(\machT\)
enjoy trace confidentiality with respect to \(R\).

@@

If, for all such pairs of traces, \(\propdef{TraceConfidentialiy} ~
\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
stack confidentiality with respect to \(\callmap\).

@@

  \item
%
    If the machine takes a \emph{call step} and $\contour$ is the contour at
    the call then,
%
    (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
    satisfied,
%
    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
    is satisfied with resepct to an arbitrary $\contour$-variant
    of $\mach$,
%
    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
    $\contour$-variant of the machine,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

@@

\apt{Need some intro text here.}

\subsection{Observable Integrity}

    \paragraph{Rolling Back to Idealized State}
      Observable integrity weakens eager integrity by allowing a returned
      state to differ from its call state at high integrity components,
      provided that the differences do not change the observable behavior
      of the rest of the program.  
%      provided the observable behavior of the induced trace from the return
%      matches that of an idealized state whose high integrity components do
%      match the call state.

      Given a contour, a call state, and a returned state, we define a
      ``rollback'' function \(\mathit{roll} : \CONTOURS \times \MACHS \times
      \MACHS \rightarrow \MACHS\). The rollback returns a state that matches
      the call state on those components that are high integrity in its
      contour, and matches the return state on low integrity components:
      \[\mathit{roll}(\contour,\mach_c,\mach_r)(\component) =
      \begin{cases}
        \mach_c(\component) & \text{if } \contour(\component) = (\HIGHINT,\_) \\
        \mach_r(\component) & \text{if } \contour(\component) = (\LOWINT,\_) \\
      \end{cases}\]

@@

      \apt{Much confusion here and below about whether $\mathit{roll}$ applies to call state
        or entry state just after.}

@@

    \definition%{Observable Trace Integrity}

      An MP-Trace enjoys {\em observable trace integrity} with respect to
      some contour \(\contour\) if its final state (if any) has observable
      behavior \apt{states don't have behavior} that prefixes that of its idealized version, constructed by
      \(\mathit{roll}\):

@@

    For our example, we return to the violation of trace integrity in
    \cref{fig:intex}, in which {\tt h} has written to {\tt z} and returned.
    Whether observable integrity holds depends on the continuing trace, shown
    in \cref{fig:intlex}. Above, we have the final state of \(\MPT\),
    \(\mach_r\), and the rolled-back state, \(\mach_r' =
    \mathit{roll}(\contour,\mach_e,\mach_r)\), in which high integrity
    components are restored to their values as of \(\mach_e\). Specifically,
    address 103 is restored to 0. When instruction 31 prints {\tt z} in the
    trace from \(\mach_r\), it prints 18, but in the trace from \(\mach_r'\)
    it prints 0. So observable trace integrity does not hold.

    However, if {\tt g} had an additional instruction to overwrite {\tt z}
    before printing, observable trace integrity would hold.
    \apt{Better to give an example in which {\tt h} overwrites something that
      is not needed to compute an output, so OTI does not fail.
      E.g., {\tt h} could overwrite {\tt w} with nonsense before returning -- but
      this doesn't matter, because the return value will replace the nonsense
      before this figures in the final print. 
      Also could be good to give an example in which we 
      overwrite {\tt x} instead of {\tt z} to show that we still fail OTI -- in other words,
      the entire remainder of the program matters, not just the caller.
    }

    \definition%{Observable Stack Integrity}

    A system enjoys {\em observable stack integrity}
    % with respect to    \(\callmap\) and \(\codemap\)
    if, for any initial state \(\mach_0\) and
      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
      and contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).
\apt{more words please}

\subsection{Observable Confidentiality}

@@

      Trace confidentiality is a property of two traces whose initial machine
      states are variations, with two sub-properties. First, the behavioral
      condition -- that their observable behaviors prior to the end of the
      trace are equivalent -- reflects the intuition that secret values should
      not influence visible behavior. Second, we have the end condition that
      their final states (if any) are equal up to the original variation. The
      end condition over-approximates the extension of the behavioral condition
      to future execution.\apt{That's quite mysterious.}

      We introduce {\em observable trace confidentiality}, which extends the
      behavioral condition to apply to execution from the final state on.
      This requires restoring varied components to their original values,
      since after return they will once again be legitimately accessible.

@@

    \definition{%Observable Trace Confidentiality}

      Let \(\MPT\) and \(\machT\) be a pair of traces starting from
      \(\mpstate{\mach_e}{\pol_e}\) and \(\mach_e'\), respectively, with
      \(\mach_e \approx_\contour \mach_e'\). \apt{Should some of this move into the defining clauses below?} Then
      $\MPT$ and $\machT$ enjoy \emph{observable trace
        confidentiality} with respect to \(\contour\), written
            \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\), if:

      \begin{itemize}
      \item
      the two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
      \pi_o(\machT)\); and

      \item
      if \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\_}\), then \(\machT\) is finite, and
      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\), where
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\),
      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \hookrightarrow \machT'\),
      and \(\mach_r' = \last(\machT)\). 
      \end{itemize}

@@

      In our example, \cref{fig:conflex} shows how execution continues from
      \(\mach_r'\) and the restored
      \[\mach_r'' = \mathit{restore}(\contour,
      \mach_e,\mach_e',\mach_r,\mach_r').
      \]
      Addresses 108 and 109
      are not restored; every address that was varied and not changed in one
      of the executions is. So 108 contains 43 in the trace from \(\mach_r\),
      and after returning from {\tt g}, {\tt main} prints 85. But in the trace
      from \(\mach_r''\), 108 contains -4, and ultimately 38 is printed.
      So observable trace confidentiality does not hold. This example
      illustrates the weakness of observable confidentiality as a testing
      property; it takes until the very end of the program to determine whether
      a tainted value actually influences observable behavior.

    \definition%{Observable Stack Confidentiality}

      A system enjoys {\em observable stack confidentiality}
      %with respect to \(\callmap\) and \(\codemap\)
      if, for any initial state
      \(\mach_0\) and each transition with entry state
      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).

    \definition%{Observable Stack Safety}
    A system enjoys {\em observable stack safety}
    % with respect to a call map       and a function map
    if it enjoys both observable stack integrity and
      observable stack confidentiality with respect to them.

\apt{Words needed to connect these ideas back to the lazy policies which motivated them in the first place!}
